---
title: "Lab 7"
author: "An Bui"
date: "2/21/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries}
library(tidyverse)
library(tmap)
library(sf)
library(spatstat)
library(maptools)
library(sp)
library(raster)
library(gstat)
```

#### 1. Raster analysis with Hawaii data in `raster` package

```{r geo tif raster information}
# hawaii
hi_par <- raster("PAR_CLIM_M.tif")
plot(hi_par)

# sst
hi_sst <- raster("SST_LTM.tif")
plot(hi_sst)

# chl A
hi_chl <- raster("CHL_LTM.tif")
plot(hi_chl)

# plot all three side by side
par(mfrow = c(1,1))
plot(hi_par)
plot(hi_sst)
plot(hi_chl)
```

##### Reproject so that this is in WGS84.

```{r reprojection}
# create a character string to reference
wgs84 <- "+proj=longlat +datum=WGS84 +ellips=WGS84 +no_defs"
# consistency with spaces is IMPORTANT

# reproject raster data
hi_sst_84 <- projectRaster(hi_sst, 
                           # coordinate reference system of choice from character string
                           crs = wgs84, 
                           # method "bilinear" for continuous data
                           method = "bilinear")
hi_sst_84@crs
# tells you that you're in WGS84

plot(hi_sst_84)
```

##### Aggregate cells by a factor of 10 using `raster::aggregate` to resample raster data.
```{r aggregate}
sst_rs <- aggregate(hi_sst, fact = 10)
plot(sst_rs)
```

##### Crop raster.
```{r crop}
hi_sst_84@extent
plot(hi_sst_84)

# make a rectangle that R recognizes as a spatial polygon
            # create an extent
bounds <- as(extent(-156.2, -154.5, 18.7, 20.4),
             # convert to spatial polygons
             'SpatialPolygons')
# hasn't been assigned a reference system

# so set a reference system
# because bounds doesn't have a crs associated with it, you can do this instead of projectRaster()
crs(bounds) <- crs(hi_sst_84)
# now bounds has the wgs84 projection but still has the projection given

# crop sst info
# crop by this raster, by this polygon
sst_crop <- crop(hi_sst_84, bounds)
plot(sst_crop)
```


##### Simple raster math

* Figure out a nonsensical variable called "tropicality" that is the sum of PAR + SST + 2*ChlA
* then map
* For thematic (categorical) raster data, could use method = "ngm" (nearest neighbor)
```{r}
hi_par_84 <- projectRaster(hi_par, crs = wgs84, method = "bilinear")

hi_chla_84 <- projectRaster(hi_chl, crs = wgs84, method = "bilinear")

plot(hi_sst_84)
plot(hi_par_84)
plot(hi_chla_84)
# plotted next to each other, scales are clearly different
```

```{r}
trop <- hi_par_84 + hi_sst_84 + 2*hi_chla_84
# gives a warning message: Raster objects have different extents. Result for their intersection is returned
# only going to return things where data has the same data in it.
```

#### Use tmap

```{r}
islands <- read_sf(dsn = 'islands', layer = "Island_boundaries") %>% 
  dplyr::select(Island) %>% 
  st_simplify(dTolerance = 10) %>% 
  st_transform(crs = 4326)

plot(islands)

# for static viewing
tmap_mode("plot")

# for interactive viewing
# tmap_mode("view")

sst_map <- tm_shape(hi_sst_84) +
  tm_raster(title = "Mean Sea Surface Temperature") +
  tm_layout(bg.color = "navyblue",
            legend.position = c("left", "bottom"),
            legend.text.color = "white",
            legend.text.size = 0.5) +
  tm_shape(islands) +
  tm_fill("darkgreen")

# create a file to store
tmap_save(sst_map, "sst_map.png", height = 5)
```

##### Conditional rasters and masking
Let's say that there is a sensitive species that needs a specific type of habitat:

* warm water (average SST >= 25.6 degrees) 
* solar irradiation (PAR) below 54
```{r}
# make extent match
extent(hi_sst_84) <- extent(hi_par_84)

# make the number of rows and columns the same
# create a raster with the same number of rows and columns and extent that I want, then fit existing data to the new raster
# make new raster match hi_par_84 information pulled from just looking at raster
cr <- raster(nrow = 822,
             ncol = 1229,
             xmn = -160.4365,
             xmx = -154.5373,
             ymn = 18.7309,
             ymx = 22.44634)

sst_new <- resample(hi_sst_84, cr, method = "bilinear")

# compareRaster() tells you if things match, or don't
compareRaster(sst_new, hi_par_84)

plot(sst_new)
plot(hi_par_84)

# make a cropped polygon ust for Kauai
bounds_main <- as(extent(-159.9, -159.2, 21.7, 22.3), 'SpatialPolygons')

crs(bounds_main) <- crs(sst_new)

par_kauai <- crop(hi_par_84, bounds_main)
sst_kauai <- crop(sst_new, bounds_main)

plot(par_kauai)
plot(sst_kauai)
```

Now only isolate regions where temp >= 25.4, PAR < 54.
```{r}
# make a copy, just in case
par_hab <- par_kauai
sst_hab <- sst_kauai

# take a subset
# set any values in par_hab >= 54 to NA values
par_hab[par_hab >= 54] <- NA
plot(par_hab)

sst_hab[sst_hab < 25.4] <- NA
plot(sst_hab)

# only plot intersections
suit_hab <- mask(sst_hab, par_hab)
plot(suit_hab)
```

#### Part 2. Point pattern analysis using red tree vole data

```{r}
# get all observations of voles from Humboldt County
voles <- read_sf(dsn = 'redtreevoledata', layer = "ds033") %>% 
  dplyr::select(COUNTY) %>% 
  filter(COUNTY == "HUM") %>% 
  # already has projection to transform
  st_transform(crs = 4326)

plot(voles)

# get Humboldt County outline from CA county shape file
humboldt <- read_sf(dsn = 'redtreevoledata', layer = "california_county_shape_file") %>% 
  filter(NAME == "Humboldt") %>% 
  dplyr::select(NAME)

# doesn't come with its own projection, so need to create crs from scratch
st_crs(humboldt) <- 4326

plot(humboldt)

tm_shape(humboldt) +
  tm_fill() +
  tm_shape(voles) +
  tm_dots(size = 0.2)

# try with ggplot
ggplot() +
  geom_sf(data = humboldt) +
  geom_sf(data = voles)
# polygon edge not found?????
```

##### Explore point patterns

* Quadrat analysis
* distance based (neighbor analysis, G-function and K-function)

```{r}
# convert voles information from sf object to spatial object to point pattern object
voles_sp <- as(voles, 'Spatial')
# class(voles_sp) : "sp"
# class(voles) : "sf"
voles_ppp <- as(voles_sp, "ppp")
# class(voles_ppp) : "ppp"

# set a bounding window of Humboldt County
humboldt_sp <- as(humboldt, 'Spatial')
humboldt_win <- as(humboldt_sp, "owin")
# class(humboldt_win) : "owin" - outer window

# combine into points and windows
voles_pb <- ppp(voles_ppp$x, voles_ppp$y, window = humboldt_win)

plot(voles_pb)
```

Try a quadrat test.
```{r}
# get voles data, choose number of columns and rows in each direction
vole_qt <- quadrat.test(voles_pb, nx = 5, ny = 10)

# null hypothesis of spatial evenness (although you will hear it called a test for "CSR") 
vole_qt
# events DO NOT reflect spatial evenness

# run both together to superimpose
plot(voles_pb)
plot(vole_qt, add = TRUE, cex = 0.5)
```

Plot kernel densities for spatial data.

```{r}
point_density <- density(voles_pb, sigma = 0.02)

plot(point_density)

# bandwidth changes will lead to different outcomes, completely dependent on what your decision is
# always report bandwidth with spatial density plot
```

```{r}
#### create a raster for voles ####
vole_raster <- raster(point_density, crs = wgs84)

tm_shape(vole_raster) +
  tm_raster()
```

Nearest neighbor:

- G-function: considers distance of each observation to nearest neighbor
    - simple mathematically, but oversimplifies how points are distributed because considers observations to nearest neighbor (and not some centralized point)
- K-function: considers how close all neighboring observations are to an event (concentric circles)

```{r g function}
#### create sequence of distances to calculate proportions ####
# just try that shit out to find the right sequence
r <- seq (0, 0.15, by = 0.005)

#### use gestimate function in envelope wrapper ####
# function = Gest
# r = r sequence
# nsim = 20 = number of Monte Carlo iterations
gfunction <- envelope(voles_pb, fun = Gest, r = r, nsim = 20)

# plot observed values as a function of r
plot(gfunction$obs ~ gfunction$r, 
     # line graph
     type = "l",
     col = "black")
# plot theoretical values as function of r
lines(gfunction$theo ~ gfunction$r,
      type = "l",
      col = "red")
# looks like data are more clustered than CSR
```

```{r k function}

```



















